$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-07-17 paper/paper.tex
Better explain the benefits of leftover typing

% Fix autoref's example treatment 
\newtheorem{example}{Example}
\newcommand{\exampleautorefname}{Example}

@@

We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{output usage context} $\Ctx_{idxs}$ to the typing rules.

This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
These leftovers can then be used as an input to another typing derivation.
Leftover typing presents three main advantages:
\begin{itemize}
  \item \textbf{Encapsulation}:
    The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgement with both an input and an output usage context.
  \item \textbf{No extensional context split}:
    No need for an extensional context splitting proof for parallel composition.
    % Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
    % This could be obtained through automated search --- but it is better yet not to need it.
    Leftover typing uses the leftovers of $P$ to type $Q$.
  \item \textbf{New theorems}:
    For the first time, theorems like \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) can be stated within the linear \picalc{} thanks to the presence of a leftover context.
\end{itemize}

@@

The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ in typing context $\gamma$ is of type $t$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
The typing judgement for processes $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.

@@

\newpage
\bibliographystyle{plainurl}
\bibliography{paper}

%%
\newpage
\appendix
\input{appendix.tex}
\end{document}

>>>>

% Fix autoref's example treatment 
\newtheorem{example}{Example}
\newcommand{\exampleautorefname}{Example}
\newcommand{\nitheoremautorefname}{Theorem}
\newcommand{\nidefinitionautorefname}{Definition}

@@

We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{(output) usage context} $\Ctx_{idxs}$ to the typing rules.
This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
These leftovers can then be used as an input to another typing derivation.

Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent explicit proofs of context splitting --- e.g., the typing rule for parallel composition would look as follows:
\begin{mathpar}
  \inferrule
  {\opctx{\Gamma}{\Delta}{\Xi} \\ \Delta \; \type{\vdash} \; P \\ \Xi \; \type{\vdash} \; Q}
  {\Gamma \; \type{\vdash} \; \comp{P}{Q}}
\end{mathpar}
Input requires a similar context split; output requires two such context splits.
This is rather cumbersome: for input and output, the proof of context split carries \emph{exactly} the same information as the typing rule for variables; for parallel composition it means that the typing derivations of both subprocesses will have no context information whatsoever until the context split is determined.

Leftover typing solves this problem by inverting the flow of information so that \textbf{the typing derivations of the subprocesses determine the context split}.
As a result, context information is threaded left to right through the typing derivation, and explicit context split proofs are no longer needed.
In addition to this, it allows a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and for the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to acquire a more general form.

@@

The \textbf{typing judgement for variables} $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ in typing context $\gamma$ is of type $t$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
The \textbf{typing judgement for processes} $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.

@@

\bibliographystyle{plainurl}
\bibliography{paper}

%%
\appendix
\input{appendix.tex}
\end{document}
